---
title: "Homework5"
author: "Steven Black"
date: "3/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework Assignment 5  (Due: March 17, 2019, midnight EST) 
 
Use the fashion_mnist data that comes with Keras to answer the questions below.  The fashion_mnist dataset has 60,000 training images and 10,000 test images. All images are 28*28 arrays and are grey scale images. The images are individual articles of clothing labeled as 0-9 as follows: 0: T-shirt/top, 1: Trouser, 2: Pullover, 3: Dress, 4: Coat, 5: Sandal, 6:  Shirt, 7: Sneaker, 8: Bag, 9: Ankle boot.  You can follow the example from last week to answer questions 1-6.

### 1.	Load the data into a variable called fashion.
```{r}

```

### 2.	Load the train_images, train_labels, test_images, and test_labels into appropriate variables.
### 3.	Build the network with one hidden layer with 512 units, and “relu” activation function.
### 4.	Compile the network with “rmsprop” as the optimizer, “categorical_crossentropy” as the loss function, and “accuracy” as the metric.

### 5.	Train the network. Remember that you will have to reshape the images and categorically encode the labels first.
### 6.	Evaluate the network using the test images.
### 7.	How many categories of clothing are there in the test set?
### 8.	What category did the model predict test image 1? 
### 9.	What is the test label for test image 1? 
### 10.	Plot the image for test image #1 and verify the result.
