---
title: "Homework10"
author: "Steven Black"
date: "4/24/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## IS 677: Introduction to Data Science Spring 2019 
Homework Assignment 10  (Due: April 28, 2019, midnight EST) 
 
Use the CIFAR10 dataset that comes with Keras. It has 50,000 training and 10,000 testing images of 10 classes: 

- airplane
- automobile
- bird
- cat
- deer
- dog
- frog
- horse
- ship
- truck 

The image size is 32\*32*3. The 3 indicates that these are color images. 

 
1.	Clear the session and load the CIFAR10 data into a variable called cifar. (5 points)
```{r}

```


2.	Create a small training dataset using the first 1000 training images (and the corresponding labels) from CIFAR10. Similarly, create a small test dataset (and the corresponding labels) using the first test 500 images from CIFAR10. (5 points).
```{r}

```


3.	Create one-hot encoding for the labels for both train and test labels. (5 points)
```{r}

```


4.	Instantiate a VGG16 convolutional base without the top layer. (5 points)
```{r}

```


5.	Extract features from the CIFAR10 images so as to fit the conv_base. (40 points)
```{r}

```


6.	Flatten the features in order to feed them to a densely connected classifier. (5 points)
```{r}

```


7.	Build a model with one dense layer with 256 units and “relu” activation, one dropout alyer with 50% dropout rate, and a dense output layer with appropriate parameters. (15 points)
```{r}

```

8.	Compile the model with categorical_crossentropy as the loss function and optimizer_rmsprop with 0.01% learning rate (lr=0.0001). (5 points)
```{r}

```

9.	Fit the model using 30 epochs. Plot the loss and accuracies. (5 points)
```{r}

```

10.	Note that the model is likely to have low accuracy. Explain why. (10 points)
