---
title: "Homework8"
author: "Steven Black"
date: "4/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## IS 677: Introduction to Data Science Spring 2019 
Homework Assignment 8  (Due: April 14, 2019, midnight EST) 
 
Use the fashion_mnist data that comes with Keras to answer the questions below.
The fashion_mnist dataset has 60,000 training images and 10,000 test images. 
All images are 28*28 arrays and are grey scale images. 
The images are individual articles of clothing labeled as 0-9 as follows: 

- 0: T-shirt/top, 
- 1: Trouser, 
- 2: Pullover, 
- 3: Dress, 
- 4: Coat, 
- 5: Sandal, 
- 6:  Shirt, 
- 7: Sneaker, 
- 8: Bag, 
- 9: Ankle boot.  

You can follow the example from last week to answer questions 1-6.

```{r}
library(keras)
```

1.	Load the data into a variable called fashion. (5 points)
```{r}
fashion <- dataset_fashion_mnist()
```

2.	Load the train_images, train_labels, test_images, and test_labels into appropriate variables. (5 points)
```{r}
```

3.	Reshape and scale the training and test images. (5 points)
```{r}
```

4.	Create one-hot encoding of the training and test labels. (5 points)
```{r}
```

5.	Build a convolutional NN for classifying the images: layer 1: 2D CONV 32 filters size 3*3, with RELU activation. Layer 2: 2D MAX_POOL with pool size=2*2, layer 3: 2D CONV 64 filters size 3*3, with RELU activation, layer 4: 2D MAX_POOL with pool size=2*2, layer 5: 2D CONV 64 filters size 3*3, layer 6: layer_flatten, layer 7: dense with 64 units and RELU activation, layer 8: you determine. (20 points)
```{r}
```

6.	Compile the model with rmsprop as the optimizer and accuracy as the metric. You decide the loss function (5 points).
```{r}
```

7.	Fit the model with 5 epochs, 0.3 validation split, and batch size of 64. (5 points)
```{r}
```

8.	Evaluate the results using the test images. (5 points)
```{r}
```

9.	Create a new CNN model to include a dropout layer with 50% dropout rate (25 points).
```{r}
```

10.	Plot the two model performances. (10 points)
```{r}
```

11.	Explain the results. (10 points)
```{r}
```

