---
title: "Homework3"
author: "Steven Black"
date: "2/18/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use the mod_diamonds dataset attached to this assignment.  The dataset is a modified version of the diamonds data you have seen before in that some of the price information is not available (NA) and the price ranges vary based on the quality of cuts.

## 1.	Load the mod_diamonds data into R.
```{r}
mod_diamonds <- read.table("mod_diammonds.txt", header = TRUE, sep = ",")
head(mod_diamonds)
```

## 2.	Write a script in R to show the number of records whose price information is not available.
```{r}
no_price_rows <- mod_diamonds[is.na(mod_diamonds$price),]
paste("There are", length(no_price_rows), "records where price information is not available. ")
```

## 3.	Find the average price of all diamonds without counting those where the price is NA.
```{r}
mean(mod_diamonds$price, na.rm = TRUE)
```

## 4.	Find the average price for each type of cut.
```{r}
cut_price <- mod_diamonds[, c("cut", "price")]
aggregate(cut_price[,2], list(cut_price$cut), mean, na.rm = TRUE)
```

## 5.	Write a script to replace the prices where the values are not available (NA) with the appropriate mean price value.  State and justify which average price you will replace the NAs with.
```{r}
head(cut_price)
```

```{r}
# install.packages("dplyr")
library(dplyr)

cut_price_replace_na <- cut_price %>%
  group_by(cut) %>% 
  mutate_if(is.numeric, function(value) ifelse(is.na(value), mean(value, na.rm = TRUE), value))
head(cut_price_replace_na)
```

As shown above the values which were previously `NA` have been replaced by the average for that group. Replacing missing values by the average is a data engineering technique which can be used to make previously unusable data amenable to data analysis and machine learning. Replacing all the missing values by the mean for all the groups would work but replacing them by the mean for their specific group yields a more accurate representation of what the data would look like if these values were not missing. 

