---
title: "Homework5"
author: "Steven Black"
date: "3/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework Assignment 5  (Due: March 17, 2019, midnight EST) 
 
Use the fashion_mnist data that comes with Keras to answer the questions below.  The fashion_mnist dataset has 60,000 training images and 10,000 test images. All images are 28*28 arrays and are grey scale images. The images are individual articles of clothing labeled as 0-9 as follows: 0: T-shirt/top, 1: Trouser, 2: Pullover, 3: Dress, 4: Coat, 5: Sandal, 6:  Shirt, 7: Sneaker, 8: Bag, 9: Ankle boot.  You can follow the example from last week to answer questions 1-6.

```{r}
library(keras)
```

### 1.	Load the data into a variable called fashion.
```{r}
fashion <- dataset_fashion_mnist()
```

### 2.	Load the train_images, train_labels, test_images, and test_labels into appropriate variables.
```{r}
train_images <- fashion$train$x
train_labels <- fashion$train$y
test_images <- fashion$test$x
test_labels <- fashion$test$y
```

### 3.	Build the network with one hidden layer with 512 units, and “relu” activation function.
```{r}
network <- keras_model_sequential() %>%
  layer_dense(units = 512, activation = "relu", input_shape = c(28 * 28)) %>%
  layer_dense(units = 10, activation = "softmax")
```

### 4.	Compile the network with “rmsprop” as the optimizer, “categorical_crossentropy” as the loss function, and “accuracy” as the metric.
```{r}
network %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
```

### 5.	Train the network. 
Remember that you will have to reshape the images and categorically encode the labels first.
```{r}
train_images_reshaped <- array_reshape(train_images, c(60000, 28 * 28))
train_images_reshaped <- train_images_reshaped / 255

test_images_reshaped <- array_reshape(test_images, c(10000, 28 * 28))
test_images_reshaped <- test_images_reshaped / 255

train_labels_categorical <- to_categorical(train_labels)
test_labels_categorical <- to_categorical(test_labels)

network %>% fit(train_images_reshaped, train_labels_categorical, epochs = 5, batch_size = 128)

```

### 6.	Evaluate the network using the test images.
```{r}
network %>% evaluate(test_images_reshaped, test_labels_categorical)
```

### 7.	How many categories of clothing are there in the test set?
```{r}
dim(test_labels_categorical)[2]
```

### 8.	What category did the model predict test image 1? 
```{r}
class_names = c('T-shirt/top',
                'Trouser',
                'Pullover',
                'Dress',
                'Coat', 
                'Sandal',
                'Shirt',
                'Sneaker',
                'Bag',
                'Ankle boot')
img_num <- 1 
predictions <- network %>% predict(test_images_reshaped)
predictions[img_num]
class_num <- apply(predictions, 1, which.max)[img_num]
paste(class_names[class_num])
```

### 9.	What is the test label for test image 1? 
```{r}
paste(class_names[test_labels[1] + 1])
```

### 10.	Plot the image for test image #1 and verify the result.
```{r}
plot(as.raster(test_images[1,,], max = 255))
```
